---
title: "Time Series Analysis of GDP and NEPSE Index"
author: "Roshan Rai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    fig_height: 6
---
<style type="text/css">

td {  /* Table  */
  font-size: 16px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 34px;
  color: Black;
  font-family: "Times New Roman", Times, serif;
  bold: TRUE;
}
h2 { /* Header 2 */
    font-size: 28px;
  color: Black;
  font-family: "Times New Roman", Times, serif;
}
h3 { /* Header 3 */
  font-size: 24px;
  font-family: "Times New Roman", Times, serif;
  color: Black;
}
body {
    color: black;
    font-size: 22px;
    font-family: 'Cambria';
    }
code.r{ /* Code block */
    font-size: 14px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

```{r message = FALSE, warning = FALSE, echo = FALSE}
options(warn = -1)
options(scipen = 10000)
options(repr.plot.width = 14.8, repr.plot.height = 7.2)
pacman::p_load(dplyr, tidyverse, ggplot2, readr, xts, forecast, tseries, TSstudio, tinytex, plotly, DT)
```

# **Time Series Analysis**

Time series analysis is a specific way of analyzing a sequence of data points collected over an interval of time. It requires a large number of data points to ensure consistency and reliability ([_Tableau.com_](https://www.tableau.com/learn/articles/time-series-analysis)). It considers that data points over a period of time may have autocorrelation or seasonal variation. The purpose of time series analysis is to understand the structure of time series data and to fit a model for forecasting. It can be used in economic forecasting, stock market analysis, sales forecasting, return projections. Different methods like Autoregression (AR), Moving Average (MR), Autoregressive Moving Average (ARMA), Autoregressive integrated Moving Average (ARIMA), Seasonal Autoregressive Integrated Moving Average (SARIMA) can be used for forecasting. Here, I will be using ARIMA model for forecasting the NEPSE index and GDP of Nepal.

## **ARIMA Model**

ARIMA stands for Autoregressive Integrated Moving Average is model for forecasting time series data by making the time series into stationary. The condition for ARIMA model is that the time series data must be stationary which can be made by differencing. A non seasonal ARIMA model is defined as **ARIMA(p,d,q)** model where; **p** is the number of autoregressive terms, **d** is the number of non seasonal differences needed for stationarity and **q** is the number of lagged forecast errors in the prediction equation. The ARIMA model is:

**Predicted Yt = Constant + Linear Combination Lags of Y (upto p lags) + Linear Combination of Lagged Forecast errors (upto q lags)**
$$Y_t = \alpha + \beta_1Y_{t-1}+\beta_2Y_{t-2}+...+\beta_pY_{1-p}\epsilon_t+\phi_1\epsilon_{t-1}+\phi_2\epsilon_{t-2}+...\phi_q\epsilon_{t-q}$$
where,<br> \(Y_{t-1}\) is the lag1 of the series, <br> \(\beta_1\) is the coefficient of autoregressive model for lag 1 that the model estimate, and \(\alpha\) is the constant.<br> Similarly, \(\epsilon_{t-1}\) is the error term for lag 1 and \(\phi_1\) is the coefficient of moving average model for lag 1.<br>

If \(d = 0: y_t = Y_t\)<br>
If \(d = 1: y_t = Y_t - Y_{t-1}\)<br>
If \(d = 2: y_t = (Y_t - Y_{t-1}) - (Y_{t-1} - Y_{t-2})\)<br>

## **NEPSE index and GDP data**

For the time series analysis, I have used the GDP data from 1960 to 2021. The data can be obtained from [_Worldbank_](https://data.worldbank.org/indicator/NY.GDP.MKTP.CD?locations=NP). For forecasting the NEPSE index, I have used the daily NEPSE index from February 2014 to Mid-September 2022. The daily NEPSE indices have been aggregated to determine the monthly average NEPSE index for the time period. The daily NEPSE index data is obtained from [_Merolagani_](https://merolagani.com/Indices.aspx). My focus was to perform the time series analysis of NEPSE index and GDP data of Nepal. However, I also wanted to look how the GDP growth rate and NEPSE annual return are related. Hence, I have also looked at the correlation coefficient between these two variables.

```{r, echo = FALSE}
#Importing datasets
nepse <- read.csv("https://raw.githubusercontent.com/Roshann-Rai/Time-series-analysis-of-GDP-and-NEPSE-data/master/nepse.csv")
gdp_growth_rate <- read.csv("https://raw.githubusercontent.com/Roshann-Rai/Time-series-analysis-of-GDP-and-NEPSE-data/master/gdp_growth_rate_np.csv")
nepse_2022_2014_daily_index <- read.csv("https://raw.githubusercontent.com/Roshann-Rai/Time-series-analysis-of-GDP-and-NEPSE-data/master/nepse_index_2022_2014.csv")
gdp <- read_csv("https://raw.githubusercontent.com/Roshann-Rai/Time-series-analysis-of-GDP-and-NEPSE-data/master/gdp.csv", show_col_types = F)
```

```{r gdp_cleaned, echo = FALSE}
#Lets clean the gdp_growth_rate dataset
gdp_cleaned <- gdp_growth_rate %>%
  rename(growth_rate = gdp) %>%             #renaming the gdp column in growth_rate
  mutate(growth_rate = as.numeric(gsub("%", '', growth_rate)))            #removing the % character and converting the growth rate column into numeric
```

```{r gdp_1997_2022, echo = FALSE}
year = c(2022)
growth_rate = c(4.1)
gdp_2022 <- data.frame(year, growth_rate)
gdp_1997_2022 <- rbind(gdp_cleaned, gdp_2022)
```

```{r nepse_2022_2014_daily_returns, echo = FALSE}
#Cleaning the nepse_2022_2014_daily_returns datasets
nepse_2022_2014 <- nepse_2022_2014_daily_index %>%
  mutate(date = as.character(date, format = "%d/%m/%Y"),
         nepse_index = as.numeric(gsub(",", "", nepse_index))) %>%     #removing , from the nepse_index column and converting into numeric format
  filter(date < "30/02/2014")      #filtering indexes before 2020/01/01
```

```{r gdp_nepal_ts, echo = FALSE}
#Wide to long
gdp_reshaped <- gather(gdp, key = "year", value = "gdp", 2:63)

#gdp of nepal
gdp_nepal <- gdp_reshaped %>%
  mutate(date = year) %>%
  select(date, gdp)

#Format date column as date
gdp_nepal$date <- as.Date(paste(gdp_nepal$date, 12, 31, sep = "/"))

#Into xts object
gdp_nepal_ts <- xts(gdp_nepal$gdp, order.by = gdp_nepal$date)
```

### **GDP of Nepal** {.tabset}
#### Plot
```{r gdp_plot, echo = FALSE}
gdp_plot <- plot_ly(gdp_nepal, x = ~date) %>%
  add_trace(y = ~gdp, type = "scatter", mode = "lines+markers", line = list(shape = "linear", color = "rgb(10,12,240)", dash = "dash"), marker = list(symbol = "circles", color = "rgb(17, 157, 255)", size = 6), connectgaps = TRUE) %>%
  layout(title = list(text = "GDP of Nepal",
                      font = list(family = "Droid Sans Mono", size = 18, xanchor = "right")),
         xaxis = list(title = "Year"),
         yaxis = list(title = "In Billions Rs.", showline = F),
         plot_bgcolor = "white")
gdp_plot
  
```

#### Table
```{r gdp_table, echo = FALSE}
gdp_reshaped %>%
  mutate(gdp_b = gdp/1000000000) %>%
  select(year, gdp_b) %>%
  datatable(
    rownames = F,
    class = "cell-border stripe",
    colnames = c("Year", "GDP in billion Rs."),
    caption = "Table 1: GDP of Nepal in Rs. Billion",
    options = list(columnDefs = list(list(className = "dt-center", targets = 0:1)))
  )
```