---
title: "Time Series Analysis of GDP and NEPSE Index"
author: "Roshan Rai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float:
      collapsed: false
    fig_height: 4.5
    fig_width: 8
---
<style type="text/css">

td {  /* Table  */
  font-size: 16px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
  bold: TRUE;
}
h1 { /* Header 1 */
  font-size: 30px;
  color: "#900C3F";
  font-family: "Times New Roman", Times, serif;
  bold: TRUE;
}
h2 { /* Header 2 */
    font-size: 24px;
  color: Black;
  font-family: "Times New Roman", Times, serif;
}
h3 { /* Header 3 */
  font-size: 20px;
  font-family: "Times New Roman", Times, serif;
  color: Black;
}
body {
    color: black;
    font-size: 18px;
    font-family: 'Cambria';
    }
code.r{ /* Code block */
    font-size: 14px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

```{r message = FALSE, warning = FALSE, echo = FALSE}
pacman::p_load(dplyr, tidyverse, ggplot2, readr, xts, forecast, tseries, TSstudio, tinytex, plotly, DT, ggplotly, patchwork)
```

# **Time Series Analysis**

Time series analysis is a specific way of analyzing a sequence of data points collected over an interval of time. It requires a large number of data points to ensure consistency and reliability ([_Tableau.com_](https://www.tableau.com/learn/articles/time-series-analysis)). It considers that data points over a period of time may have autocorrelation or seasonal variation. The purpose of time series analysis is to understand the structure of time series data and to fit a model for forecasting. It can be used in economic forecasting, stock market analysis, sales forecasting, return projections. Different methods like Autoregression (AR), Moving Average (MR), Autoregressive Moving Average (ARMA), Autoregressive integrated Moving Average (ARIMA), Seasonal Autoregressive Integrated Moving Average (SARIMA) can be used for forecasting. Here, I will be using ARIMA model for forecasting the NEPSE index and GDP of Nepal.

## **ARIMA Model**

ARIMA stands for Autoregressive Integrated Moving Average is model for forecasting time series data by making the time series into stationary. The condition for ARIMA model is that the time series data must be stationary which can be made by differencing. A non seasonal ARIMA model is defined as **ARIMA(p,d,q)** model where; **p** is the number of autoregressive terms, **d** is the number of non seasonal differences needed for stationarity and **q** is the number of lagged forecast errors in the prediction equation. The ARIMA model is:

**Predicted Yt = Constant + Linear Combination Lags of Y (upto p lags) + Linear Combination of Lagged Forecast errors (upto q lags)**
$$Y_t = \alpha + \beta_1Y_{t-1}+\beta_2Y_{t-2}+...+\beta_pY_{1-p}\epsilon_t+\phi_1\epsilon_{t-1}+\phi_2\epsilon_{t-2}+...\phi_q\epsilon_{t-q}$$
where,<br> \(Y_{t-1}\) is the lag1 of the series, <br> \(\beta_1\) is the coefficient of autoregressive model for lag 1 that the model estimate, and \(\alpha\) is the constant.<br> Similarly, \(\epsilon_{t-1}\) is the error term for lag 1 and \(\phi_1\) is the coefficient of moving average model for lag 1.<br>

If \(d = 0: y_t = Y_t\)<br>
If \(d = 1: y_t = Y_t - Y_{t-1}\)<br>
If \(d = 2: y_t = (Y_t - Y_{t-1}) - (Y_{t-1} - Y_{t-2})\)<br>

## **NEPSE index and GDP data**

For the time series analysis, I have used the GDP data from 1960 to 2021. The data can be obtained from [_Worldbank_](https://data.worldbank.org/indicator/NY.GDP.MKTP.CD?locations=NP). For forecasting the NEPSE index, I have used the daily NEPSE index from February 2014 to Mid-September 2022. The daily NEPSE indices have been aggregated to determine the monthly average NEPSE index for the time period. The daily NEPSE index data is obtained from [_Merolagani_](https://merolagani.com/Indices.aspx). My focus was to perform the time series analysis of NEPSE index and GDP data of Nepal. However, I also wanted to look how the GDP growth rate and NEPSE annual return are related. Hence, I have also looked at the correlation coefficient between these two variables.

```{r, echo = FALSE}
#Importing datasets
nepse <- read.csv("https://raw.githubusercontent.com/Roshann-Rai/Time-series-analysis-of-GDP-and-NEPSE-data/master/nepse.csv")
gdp_growth_rate <- read.csv("https://raw.githubusercontent.com/Roshann-Rai/Time-series-analysis-of-GDP-and-NEPSE-data/master/gdp_growth_rate_np.csv")
nepse_2022_2014_daily_index <- read.csv("https://raw.githubusercontent.com/Roshann-Rai/Time-series-analysis-of-GDP-and-NEPSE-data/master/nepse_index_2022_2014.csv")
gdp <- read_csv("https://raw.githubusercontent.com/Roshann-Rai/Time-series-analysis-of-GDP-and-NEPSE-data/master/gdp.csv", show_col_types = F)
```

```{r gdp_cleaned, echo = FALSE}
#Lets clean the gdp_growth_rate dataset
gdp_cleaned <- gdp_growth_rate %>%
  rename(growth_rate = gdp) %>%             #renaming the gdp column in growth_rate
  mutate(growth_rate = as.numeric(gsub("%", '', growth_rate)))            #removing the % character and converting the growth rate column into numeric
```

```{r gdp_1997_2022, echo = FALSE}
year = c(2022)
growth_rate = c(4.1)
gdp_2022 <- data.frame(year, growth_rate)
gdp_1997_2022 <- rbind(gdp_cleaned, gdp_2022)
```

```{r nepse_2022_2014_daily_returns, echo = FALSE}
#Cleaning the nepse_2022_2014_daily_returns datasets
nepse_2022_2014 <- nepse_2022_2014_daily_index %>%
  mutate(date = as.character(date, format = "%d/%m/%Y"),
         nepse_index = as.numeric(gsub(",", "", nepse_index))) %>%     #removing , from the nepse_index column and converting into numeric format
  filter(date < "30/02/2014")      #filtering indexes before 2020/01/01
```

```{r wide to long, echo = FALSE}
#Wide to long
gdp_reshaped <- gather(gdp, key = "year", value = "gdp", 2:63)

#gdp of nepal
gdp_nepal <- gdp_reshaped %>%
  mutate(date = year) %>%
  select(date, gdp)

#Format date column as date
gdp_nepal$date <- as.Date(paste(gdp_nepal$date, 12, 31, sep = "/"))

#Into xts object
gdp_nepal_ts <- xts(gdp_nepal$gdp, order.by = gdp_nepal$date)
```

### **GDP of Nepal** {.tabset}
#### Plot
```{r gdp_plot, echo = FALSE}
# gdp_plot
gdp_plot <- gdp_reshaped %>%
  select(year, gdp) %>%
  mutate(year = as.numeric(year),
         gdp = gdp/1000000000) %>%
  ggplot(aes(year, gdp)) +
 geom_point(size = 1.5, alpha = 0.8, color = "#1930A6") +
  geom_line(linetype = 1, size = 0.9, alpha = 0.9, color = "#AEB8EE") +
  labs(title = "GDP of Nepal",
       subtitle = "From 1960 - 2021",
       x = "",
       y = "GDP in Billions Rs.",
       caption = "Source: World Bank") +
  scale_x_continuous(breaks = c(1960, 1970, 1980, 1990, 2000, 2010, 2020)) +
  theme_bw(base_size = 15)
ggplotly(gdp_plot)
```

#### Table
```{r gdp_table, echo = FALSE}
gdp_reshaped %>%
  mutate(gdp_b = gdp/1000000000) %>%
  select(year, gdp_b) %>%
  datatable(
    rownames = F,
    class = "cell-border stripe",
    colnames = c("Year", "GDP in billion Rs."),
    caption = "Table 1: GDP of Nepal in Rs. Billion",
    options = list(columnDefs = list(list(className = "dt-center", targets = 0:1)))
  )
```

### **NEPSE Index** {.tabset}

```{r xts objects for Nepse, echo = FALSE}
#Converting the nepse_2022_2019 dataframe in xts 
nepse_2022_2014_ts <- xts(nepse_2022_2014$nepse_index, order.by = as.POSIXct(nepse_2022_2014$date, format = "%d/%m/%Y"))
```

```{r monthly return, echo = FALSE}
#Monthly return
nepse_2014_2022_monthly <- apply.monthly(nepse_2022_2014_ts, FUN = mean)
```

```{r yearly avg return, echo = FALSE}
#Extract nepse index for 2015
nepse_2015_ts <- nepse_2022_2014_ts["2015",]
nepse_2015 <- round(mean(apply.yearly(nepse_2015_ts, FUN = mean)), digit = 2)

#Extract nepse index for 2016
nepse_2016_ts <- nepse_2022_2014_ts["2016",]
nepse_2016 <- round(mean(apply.yearly(nepse_2016_ts, FUN = mean)), digit = 2)

#Extract nepse index for 2017
nepse_2017_ts <- nepse_2022_2014_ts["2017",]
nepse_2017 <- round(mean(apply.yearly(nepse_2022_2014_ts, FUN = mean)), digit = 2)

#Extract nepse index for 2018
nepse_2018_ts <- nepse_2022_2014_ts["2018",]
nepse_2018 <- round(mean(apply.yearly(nepse_2018_ts, FUN = mean)), digit = 2)

#Extract nepse index for 2019
nepse_2019_ts <- nepse_2022_2014_ts["2019",]
nepse_2019 <- round(mean(apply.yearly(nepse_2019_ts, FUN = mean)), digit = 2)

#Extract nepse index for 2020
nepse_2020_ts <- nepse_2022_2014_ts["2020",]
nepse_2020 <- round(mean(apply.yearly(nepse_2020_ts, FUN = mean)), digit = 2)

#Extract nepse index for 2021
nepse_2021_ts <- nepse_2022_2014_ts["2021",]
nepse_2021 <- round(mean(apply.yearly(nepse_2021_ts, FUN = mean)), digit = 2)

#Extract nepse index for 2022
nepse_2022_ts <- nepse_2022_2014_ts["2022",]
nepse_2022 <- round(mean(apply.yearly(nepse_2022_ts, FUN = mean)), digit = 2)
```

```{r nepse_2015_2022, echo = F}
#Creating a new dataframe
year <- c("2015", "2016", "2017", "2018", "2019", "2020", "2021", "2022")
nepse_index <- c("1028.88", "1539.43", "1553.74", "1261.82", "1192.81", "1523.69", "2700.05", "2314.07")
nepse_2015_2022 <- data.frame(year, nepse_index)

#Format year column as date
nepse_2015_2022$year <- as.Date(paste(nepse_2015_2022$year, 12, 31, sep = "/"))

#Creating time series object
nepse_2015_2022_ts <- xts(as.numeric(nepse_2015_2022$nepse_index), order.by = nepse_2015_2022$year)

#Converting year column of nepse dataframe into date format
nepse$year <- as.Date(paste(nepse$year, 12, 31, sep = "/"))

#converting nepse dataframe into xts object
nepse_1997_2019_ts <- xts(nepse$nepse_index, order.by = nepse$year)
nepse_1997_2014_ts <- nepse_1997_2019_ts["1997/2014",]

#Merging the xts objects
nepse_ts <- rbind(nepse_1997_2014_ts, nepse_2015_2022_ts)
```

#### Plot
```{r nepse_df, echo = F}
#Creating gdp_nepse_df
nepse_df <- as.data.frame(nepse_ts)

#Naming the column date
nepse_df$year <- rownames(nepse_df)    # Convert row names to column
rownames(nepse_df) <- NULL           # Reset row names

#plotting nepse_df
nepse_plot <- nepse_df %>%
  mutate(index = V1,
         year = as.Date(year)) %>%
  ggplot(aes(year, index)) +
 geom_point(size = 1.5, alpha = 0.8, color = "#E00D19") +
  geom_line(linetype = 1, size = 0.9, alpha = 0.9, color = "#FCA3A8") +
  labs(title = "NEPSE Index",
       subtitle = "From 1997 - 2022",
       x = "",
       y = "Index",
       caption = "Source: Merolagani") +
  scale_y_continuous(breaks = c(500, 1000, 1500, 2000, 2500, 3000)) +
  theme_bw(base_size = 15)
ggplotly(nepse_plot)
```

#### Table
```{r nepse_table, echo = F}
nepse_df %>%
  mutate(index = V1) %>%
  select(year, index) %>%
  datatable(
    rownames = F,
    class = "cell-border stripe",
    colnames = c("Year", "Index"),
    caption = "Table 2: NEPSE Yearly Index",
    options = list(columnDefs = list(list(className = "dt-center", targets = 0:1)))
  )
```

### **GDP growth and NEPSE return** {.tabset}
```{r, yearly_return, echo = F}
#Yearly return on nepse 
return_annual <- (round(diff(log(nepse_ts))[-1,], digit = 4))*100
```

#### Plot
```{r, gdp_nepse xts, echo = F}
#Formatting year column of gdp_1997_2022 as character
gdp_1997_2022$year <- as.character(gdp_1997_2022$year)

#Formatting year column of gdp_1997_2022 as date
gdp_1997_2022$year <- paste(gdp_1997_2022$year, 12, 31, sep = "/")
gdp_ts <- xts(gdp_1997_2022$growth_rate, order.by = as.Date(gdp_1997_2022$year))


#merge gdp and nepse xts objects
nepse_gdp_ts <- merge.xts(gdp_ts, return_annual)[-1,]

#Creating gdp_nepse_return_df
gdp_nepse_return_df <- as.data.frame(nepse_gdp_ts)

#Naming the column date
gdp_nepse_return_df$year <- rownames(gdp_nepse_return_df)    # Convert row names to column
rownames(gdp_nepse_return_df) <- NULL           # Reset row names

#Plotting gdp growth rate and nepse return
return_rate <- gdp_nepse_return_df %>%
  mutate(gdp_growth_rate = gdp_ts,
         nepse_annual_return = return_annual,
         year = as.Date(year)) %>%
  select(year, gdp_growth_rate, nepse_annual_return)

return_rate_plot <- ggplot() +
  geom_point(return_rate, mapping = aes(x = year, y = gdp_growth_rate), color = "#1930A6", alpha = 0.9) +
  geom_point(return_rate, mapping = aes(x = year, y = nepse_annual_return), color = "#E00D19", alpha = 0.9) +
  geom_line(return_rate, mapping = aes(x = year, y = gdp_growth_rate), color = "#AEB8EE", size = 0.7, alpha = 0.6) +
  geom_line(return_rate, mapping = aes(x = year, y = nepse_annual_return), color = "#FCA3A8", size = 0.7, alpha = 0.6) +
  labs(title = "GDP growth rate and NEPSE annual Return",
       y = "In Percentage",
       x = "") +
  scale_x_date(date_labels = "%Y", date_breaks = "5 years") +
  theme_bw(base_size = 14)

ggplotly(return_rate_plot)
```

#### Table
```{r gdp_growth and nepse return, echo = F}
return_rate %>%
    datatable(
    rownames = F,
    class = "cell-border stripe",
    colnames = c("Year", "GDP growth rate", "NEPSE Annual Return"),
    caption = "Table 3: GDP Growth Rate and NEPSE Index Returns",
    options = list(columnDefs = list(list(className = "dt-center", targets = 0:1)))
  )
```

### **Correlation Coefficient between GDP growth rate and NEPSE annual return**
Correlation Coefficient measures the strength of relationship between two variables. The magnitude of correlation coefficient ranges from -1.0 to +1.0. A correlation coefficient of -1.0 is perfect negative correlation i.e. when one variable increases by 1 unit other decreases by 1 unit. Similarly, a correlation coefficient of +1.0 is perfect positive correlation coefficient i.e. when one variable increases by 1 unit other variable increases by 1 unit. Karl Pearson Correlation Coefficient is given by:
$$r = \frac{\sum(X-\bar{X})(Y-\bar{Y})}{{\sqrt{\sum(X-\bar{X})^2}}{\sqrt{\sum(Y-\bar{Y})^2}}}$$
Using the Karl Pearson correlation coefficient on the data from 1998 to 2022, the correlation coefficient between **NEPSE Annual Return** and **GDP growth rate** is found to be -0.0077. This means insignificant (very weak) negative relationship between those two variables.
```{r, echo =F}
cor_return_rate <- return_rate %>%
  select(gdp_growth_rate, nepse_annual_return)
cor <- cor(cor_return_rate, method = "pearson")
```

## **Forecasting**{.tabset}
### **Forecasting GDP**
I have used ARIMA model to forecast the GDP of Nepal. For forecasting, I have used the GDP data from 1960 - 2021.

#### **Test of Stationarity**
A stationary time series is one whose mean and variance don't change with time. Thus, time series with trends, or with seasonality, are not stationary. While, a white noise series is stationary ([_Towardsdatascience_](https://towardsdatascience.com/stationarity-assumption-in-time-series-data-67ec93d0f2f)).
```{r, echo = F, class.output = "bg.warning"}
#Test of Stationarity
adf.test(gdp_nepal_ts, k = 15)
```
At lag = 15, p-value from Augmented Dickey-Fuller test is 0.99 which is greater than \(\alpha\)=0.05 at 95% confidence interval. This suggests that null hypothesis is accepted i.e. the time series data is not stationary.
```{r, echo = F, class.source = "bg-danger", fig.height = 4, fig.width = 6}
#Forecasting GDP
#Plotting acf and pacf
acf(gdp_nepal_ts, main = "ACF", xlim = c(1,20)) 
pacf(gdp_nepal_ts, main = "PACF", xlim = c(1,20)) 
```

Looking at the **acf** and **pacf** plot, we can see the vertical lines crossing the **blue** horizontal line is substantial. This supports the **adf test**. 

#### **ARIMA Model**

Using the Akaike Information Criterion (AIC) for evaluating the ARIMA model, the best ARIMA model is ARIMA(2,2,1). AIC is the mathematical method for evaluating how well a model fits the data it was generated from. Among the ARIMA models generated, ARIMA(2,2,1) has the lowest AIC value i.e. 2662.796. In this model p = 2 which is the AR term, d = 2 is the number of differencing needed and q = 1 is the MA term.The coefficients of AR1, AR2 and MA1 are 0.0781, -0.3977 and -0.7538 respectively.
```{r,  echo = F, class.source = "bg-danger", class.output = "bg.warning"}
#Creating the model using auto.arima()
model <- auto.arima(gdp_nepal_ts, ic="aic", trace = T)
model
```

#### **Test of Stationarity of the model**
Plotting the **acf** and **pacf** plots of the residuals of the model show that the number of verticle lines crossing the horizontal blue lines have reduced significantly indicating the stationarity of the time series model.
```{r, echo = F, class.source = "bg-danger", class.output = "bg.warning", fig.height = 4, fig.width = 6}
#Has the model been stationary?
acf(ts(model$residuals), xlim = c(1,20))
pacf(ts(model$residuals), xlim = c(1,20))
```

#### **Forecasting**
Using the ARIMA(2,2,1) model, I have made the forecasting of GDP for next 10 years at confidence interval of 95%.
```{r, echo =F}
#Lets forecast the gdp for next 10 years at 95% confidence interval
gdp_forecast <- forecast(model, level = c(95), h = 10)

#Creating the dataframe for predictions
year <- c(seq(2022, 2031, by=1))
Point_forecast <- c(39.12, 40.91, 42.34, 44.31, 46.40, 48.28, 50.09, 51.99, 53.91, 55.82)
Low_95 <- c(37.36, 37.68, 38.35, 39.58, 40.73, 41.57, 42.35, 43.19, 43.99, 44.73)
High_95 <- c(41.26, 44.15, 46.32, 49.04, 52.08, 54.99, 57.84, 60.79, 63.84, 66.91)
gdp_forecast_10 <- data.frame(year, Point_forecast, Low_95, High_95)

#Creating table
gdp_forecast_10 %>%
    datatable(
    rownames = F,
    class = "cell-border stripe",
    colnames = c("Year", "Point Forecast", "Low 95%", "High 95%"),
    caption = "Table 4: Forecasted GDP in Billions Rs.",
    options = list(columnDefs = list(list(className = "dt-center", targets = 0:1)))
  )
plot(gdp_forecast)
```

### **Forecasting NEPSE Index**
For forecasting the NEPSE index, I have used the daily NEPSE index data from February 2014 to September 2022 which is aggregated for average monthly index during the period. The NEPSE index has been forecasted using the ARIMA model similar to above.

#### **Test of Stationarity**
The Augmented Dickey-Fuller test at the lag of 15 gives the p-value 0.429 which is much greater than \(\alpha\)=0.05 at 95% confidence interval. Hence, it supports that the null hypothesis is accepted i.e. the time series data is stationarity.<br>
Similarly, **acf** and **pacf** plots have significant number of vertical lines crossing the blue horizontal line. This also supports the result of adf test.
```{r, echo = F}
#Testing stationarity
adf.test(nepse_2014_2022_monthly, k = 15)

#Plotting acf and pacf
par(mfrow = c(1,2))
acf(nepse_2014_2022_monthly)
pacf(nepse_2014_2022_monthly)
```

#### **Model**
Using the **auto.arima()** function, the best model is ARIMA(2,1,2) for which the AIC is 1255.602. For this model p = 2 is the AR term, d = 1 is the number of differencing required and q = 2 is the MA term. The coefficients of AR1, AR2, MA1 and MA2 are -0.0734, -0.604, 0.1496 and 0.8870.
```{r, echo = F}
#Creating the model using auto.arima()
model_nepse <- auto.arima(nepse_2014_2022_monthly, ic="aic", trace = T)
model_nepse
```

#### **Test of Stationarity test for model's residuals**
The **acf** and **pacf** plots have very minimum number of vertical lines crossing the blue horizontal lines indicating the stationarity of the model.
```{r, echo = F}
#Has the model been stationary?
par(mfrow = c(1,2))
acf(ts(model_nepse$residuals))
pacf(ts(model_nepse$residuals))
```

#### **Forecasting**
For forecasting the NEPSE index, I have used the daily NEPSE index from February 2014 to September 2022 from which average monthly NEPSE index has been calculated. Similar to GDP forecasting, I have used the ARIMA model and forecasted the NEPSE index for next 12 months.

```{r, echo = F}
#Lets forecast the gdp for next 10 years at 95% confidence interval
nepse_forecast <- forecast(model_nepse, level = c(95), h = 12)

#Creating the dataframe for predictions
Month <- c("2022 October", "2022 November", "2022 December", "2023 January", "2023 February", "2023 March", "2023 April", "2023 May", "2023 June", "2023 July", "2023 August", "2023 September")
Point_forecast <- c(1974.362, 1923.131, 1909.147, 1941.116, 1947.215, 1927.458, 1925.225, 1937.322, 1937.782, 1930.442, 1930.703, 1935.118)
Low_95 <- c(1758.375, 1605.838, 1477.696, 1427.844, 1379.376, 1305.349, 1245.239, 1207.073, 1164.448, 1114.447, 1072.126, 1037.162)
High_95 <- c(2190.350, 2240.424, 2340.597, 2454.389, 2515.054, 2549.566, 2605.212, 2667.572, 2711.117, 2746.437, 2789.280, 2833.073)
nepse_forecast_10 <- data.frame(Month, Point_forecast, Low_95, High_95)

# #Creating table
nepse_forecast_10 %>%
    datatable(
    rownames = F,
    class = "cell-border stripe",
    colnames = c("Month", "Point Forecast", "Low 95%", "High 95%"),
    caption = "Table 5: Forecasted NEPSE Index.",
    options = list(columnDefs = list(list(className = "dt-center", targets = 0:1)))
  )

plot(nepse_forecast)
```
## **Conclusion**
ARIMA is found to be more effective especially for short term time series forecasting (Box 1970; Jarrett 1991). ARIMA model can increase the effectiveness of forecasting even with the minimum parameters. However, long term forecasting eventually goes to be straight line and it is poor at forecasting series with turning points.<br>
This blog explains the basic of ARIMA model and presents the process to perform the time series forecasting using this model. For more information regarding ARIMA model, visit [_here_](https://otexts.com/fpp2/arima.html). My code for time series analysis can be found [_here_](https://github.com/Roshann-Rai/Time-series-analysis-of-GDP-and-NEPSE-data/blob/master/gdp_nepse_time_series_analysis.R).<br>
Lastly all thanks to [_Code For Nepal_](https://codefornepal.org/) for providing me with opportunity to learn R.
